---
title: 'Case Study - Virginia Counties'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case Study - Virginia Counties}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

_Built with R
`r getRversion()`_

---

```{css, echo = FALSE}
.table {display: table}
.table th {text-align: right}
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height = 5,
  fig.width = 8.84,
  dev = "CairoSVG",
  fig.ext = "svg"
)
dir <- "../dc.hifld.ems/data/working/"
if (!dir.exists(dir)) dir <- "../../dc.hifld.ems/data/working/"
library(sf)
library(Matrix)
library(catchment)
library(splot)
library(leaflet)
library(knitr)
```

This case study compares different ways of calculating catchment area ratios between two counties in Virginia:
Loudoun and Fairfax. These counties represent more rural (Loudoun) and more urban (Fairfax) environments.

# Data

**Demand** is defined as the total population in 2020, as estimated in the [U.S. Census Bureau](https://www.census.gov)'s
[American Community Survey](https://www.census.gov/programs-surveys/acs) 5-year summaries.

**Supply** is defined as hospital beds, as collected from various government sources by Oak Ridge National Laboratory, and hosted by [Homeland Infrastructure Foundation-Level Data](https://hifld-geoplatform.opendata.arcgis.com): [Hospitals](https://hifld-geoplatform.opendata.arcgis.com/datasets/geoplatform::hospitals/about). The inclusion criteria for hospitals were that they have a status of `OPEN` and a type of either `CRITICAL ACCESS` or `GENERAL ACCUTE CARE`. Missing bed counts (13 of 805 within the working subset) were median-imputed.

**Cost** is defined as travel times between each region's centroid and each hospital, as calculated within road networks by the [Open Source Routing Machine](https://github.com/Project-OSRM/osrm-backend) using [OpenStreetMap](https://www.openstreetmap.org) data from 2022, as provided by [Geofabrik](https://download.geofabrik.de/north-america/us.html).

Data were prepared as part of a data commons project; the full script is available at [uva-bi-sdad/dc.hifld.ems/src/retrieve.R](https://github.com/uva-bi-sdad/dc.hifld.ems/blob/main/src/retrieve.R).
Here, we will be loading in the files created from the original sources.

## Downloading/Loading

You can clone the repository, or directly download the files in the [data/working](https://github.com/uva-bi-sdad/dc.hifld.ems/tree/main/data/working) directory, then point to the directory containing those files with the `dir` variable:

```sh
git clone --branch main https://github.com/uva-bi-sdad/dc.hifld.ems.git
```

```r
# this would be the directory when working from within the dc.hifld.ems repository
dir <- "data/working/"
```

Now, we can load the files, which will results in 3 main objects: `hospitals`, `population`, and `traveltimes`:

```{r}
# main objects
hospitals <- read.csv(gzfile(paste0(dir, "hospitals.csv.xz")))
population <- read.csv(gzfile(paste0(dir, "population_bg.csv.xz")))
rownames(population) <- population$GEOID
population$tract <- substring(population$GEOID, 1, 11)
traveltimes <- do.call(rbind, lapply(
  paste0(dir, "traveltimes_2020_", unique(substring(population$GEOID, 1, 2)), ".csv.xz"),
  function(f) read.csv(gzfile(f), row.names = 1, check.names = FALSE)
))

# tract-level comparisons
population_tract <- read.csv(gzfile(paste0(dir, "population_tr.csv.xz")))
rownames(population_tract) <- population_tract$GEOID
population_tract <- population_tract[unique(population$tract), ]
traveltimes_tract <- do.call(rbind, lapply(
  paste0(dir, "traveltimes_2020_", unique(substring(population_tract$GEOID, 1, 2)), "_tract.csv.xz"),
  function(f) read.csv(gzfile(f), row.names = 1, check.names = FALSE)
))
```

We'll also use a map to show comparisons between the focal regions:

```{r}
# download map geographies
library(sf)
tracts <- st_read(paste0(
  "https://raw.githubusercontent.com/uva-bi-sdad/dc.geographies/main/data/",
  "va_geo_census_cb_2020_census_tracts/distribution/",
  "va_geo_census_cb_2020_census_tracts.geojson"
), quiet = TRUE)
tracts <- tracts[grep("^51(?:107|059|600)", tracts$geoid), ]
block_groups <- st_read(paste0(
  "https://raw.githubusercontent.com/uva-bi-sdad/dc.geographies/main/data/",
  "va_geo_census_cb_2020_census_block_groups/distribution/",
  "va_geo_census_cb_2020_census_block_groups.geojson"
), quiet = TRUE)
block_groups <- block_groups[grep("^51(?:107|059|600)", block_groups$geoid), ]

# make base maps
library(leaflet)
map_tracts <- leaflet(tracts, options = leafletOptions(attributionControl = FALSE)) |>
  setView(-77.4, 38.99, 9) |>
  addProviderTiles("CartoDB.Positron") |>
  addScaleBar("bottomleft") |>
  addMapPane("lines", zIndex = 410) |>
  addMapPane("points", zIndex = 411) |>
  addLayersControl(position = "topleft", overlayGroups = "Hospitals") |>
  addCircles(
    data = hospitals, color = "#000", fillColor = "#000", opacity = .8, lng = ~X, lat = ~Y,
    label = ~ paste0("ID: ", ID, ", Beds: ", BEDS),
    group = "Hospitals", options = pathOptions(pane = "points")
  )
map_block_groups <- leaflet(block_groups, options = leafletOptions(attributionControl = FALSE)) |>
  setView(-77.4, 38.99, 9) |>
  addProviderTiles("CartoDB.Positron") |>
  addScaleBar("bottomleft") |>
  addMapPane("lines", zIndex = 410) |>
  addMapPane("points", zIndex = 411) |>
  addLayersControl(position = "topleft", overlayGroups = "Hospitals") |>
  addCircles(
    data = hospitals, color = "#000", fillColor = "#000", opacity = .8, lng = ~X, lat = ~Y,
    label = ~ paste0("ID: ", ID, ", Beds: ", BEDS),
    group = "Hospitals", options = pathOptions(pane = "points")
  )

# specify colors
palette <- scico::scico(255, direction = -1, palette = "vik")
plot_colors <- c("#b69349", "#003e7d")
```

We'll be looking at several variants of floating catchment areas, so, to better compare them,
we'll start with a baseline that represents the most thorough and up-to-date variant:

```{r}
library(catchment)
baseline <- catchment_ratio(
  population, hospitals, traveltimes,
  weight = "gaussian", scale = 18, normalize_weight = TRUE, return_type = 1e5,
  consumers_value = "population", providers_id = "ID", providers_value = "BEDS"
)
names(baseline) <- population$GEOID
```

# Resolution

In all of our catchment area calculations, we are referring to resources per person, but we do not
calculate costs for each individual. In this respect, our calculations are always approximations.

To get a feel for how much impact this approximation has, we can compare cost calculations made at
the highest resolution with population data available (block groups; for yearly estimates) with
those made at a lower resolution (tracts).

```{r}
# calculate floating catchment area ratios
resolutions <- data.frame(
  # starting from block groups and aggregating up to tracts
  higher = vapply(split(data.frame(
    ratio = baseline,
    population = population$population
  ), population$tract), function(bg) {
    total <- sum(bg$population)
    if (total) sum(bg$ratio * bg$population) / total else 0
  }, 0),
  # starting from tracts
  lower = catchment_ratio(
    population_tract, hospitals, traveltimes_tract,
    weight = "gaussian", scale = 18, normalize_weight = TRUE, return_type = 1e5,
    consumers_value = "population", providers_id = "ID", providers_value = "BEDS"
  )
)
```

Overall, calculations made at higher and lower resolutions have a Person's _r_ of `r round(cor(resolutions$higher, resolutions$lower), 3)`.

Looking specifically at our comparison counties...

```{r}
# convenience functions to look at the focal counties
library(knitr)
get_focal_subset <- function(data, rowIds, counties = c(Loudoun = "^51107", Fairfax = "^51(?:600|059)")) {
  comp <- colnames(data)[1:2]
  sub <- do.call(rbind, unname(lapply(counties, function(county) {
    d <- data[grep(county, rownames(data)), ]
    n <- nrow(d)
    d[[paste0("rank_", comp[[1]])]] <- order(-d[, 1]) / n * 100
    d[[paste0("rank_", comp[[2]])]] <- order(-d[, 2]) / n * 100
    d$County <- names(counties[counties == county])
    d$difference <- d[, 1] - d[, 2]
    d$rank_difference <- d[[paste0("rank_", comp[[1]])]] - d[[paste0("rank_", comp[[2]])]]
    d
  })))
  data <- sub[rowIds, ]
  list(data = data, pal = colorNumeric(palette, rep(max(abs(data$difference)), 2) * c(-1, 1)))
}
summary_contrast <- function(subset) {
  data <- subset$data
  summary <- vapply(split(data, data$County), function(d) {
    c(
      colMeans(d[, 1:2]), mean(d$difference), cor(d[, 1], d[, 2]), mean(abs(d$rank_difference))
    )
  }, numeric(5))
  rownames(summary) <- c(
    paste0("Average (", colnames(data)[1:2], ")"),
    "Average Difference", "Correlation", "Average Change in Rank Percent"
  )
  kable(summary, digits = 3)
}
```

```{r, result='asis'}
res_subset <- get_focal_subset(resolutions, tracts$geoid)
summary_contrast(res_subset)
```
Average difference is in number of beds per 100k people; higher means more are estimated when starting from tracts.

```{r}
map_tracts |>
  addControl("Difference in Hospital Beds Per 100K People (Block Groups - Tract)", "topright") |>
  addLegend("bottomright", res_subset$pal, res_subset$data$difference, opacity = 1) |>
  addPolygons(
    fillColor = res_subset$pal(res_subset$data$difference), fillOpacity = 1, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Beds",
    label = paste0("Block Groups - Tract: ", round(res_subset$data$difference, 3))
  )
```

Higher values (blue) mean ratios are higher when made at the block group level.

```{r}
library(splot)
splot(
  higher ~ lower * County, res_subset$data,
  colors = plot_colors,
  title = FALSE, laby = "From Block Groups", labx = "From Tracts"
)
```

Ratios are even more highly correlated within county.

The only reason to use lower resolution geographies would be if (a) population data are not available
at higher resolutions, or (b) there are practical computing limits.

# Distance

Our next comparison will consider different means of calculating distance / travel time, which
represents cost, and define the shape of the catchment areas.

The simplest means of calculating distance would be to measure straight lines between each consumer
and provider location (resulting in perfectly circular catchment areas).

We use a routing machine to get estimated travel times within road networks (resulting in catchment
areas shaped by the local road topology), which should be more accurate, but requires some setup
and computing resources.

```{r}
distances <- data.frame(
  # with cost defined by travel times as routed within road networks
  routed = baseline,
  # with cost defined by Euclidean distances
  straight = catchment_ratio(
    population, hospitals,
    weight = "gaussian", scale = .13, normalize_weight = TRUE, return_type = 1e5,
    consumers_value = "population", providers_id = "ID", providers_value = "BEDS"
  )
)
```

Overall, calculations made with routed and straight costs have a Person's _r_ of `r round(cor(distances$routed, distances$straight), 3)`,
though this is fairly sensitive to the weight function.

Looking specifically at our comparison counties...

```{r, result='asis'}
dist_split <- get_focal_subset(distances, block_groups$geoid)
summary_contrast(dist_split)
```

```{r}
map_block_groups |>
  addControl("Difference in Hospital Beds Per 100K People (Routed - Straight)", "topright") |>
  addLegend("bottomright", dist_split$pal, dist_split$data$difference, opacity = 1) |>
  addPolygons(
    fillColor = dist_split$pal(dist_split$data$difference), fillOpacity = 1, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Beds",
    label = paste0("Routed - Straight: ", round(dist_split$data$difference, 3))
  )
```

Higher values (blue) mean ratios are higher when costs are based on routed travel time.

```{r}
splot(
  routed ~ straight * County, dist_split$data,
  colors = plot_colors,
  title = FALSE, laby = "With Route Times", labx = "With Straight Distance"
)
```

Correlations within subsets remain high, with some interesting skewing.

The only reason to use straight distances in this context would be if good travel time
data are not available.

# Buffer

There is a warping effect at the boundaries of catchment area calculation regions if the regions
are artificially restricted, such as at state borders when consumers can freely access out-of-state
providers. To mitigate such a warping effect, we include surrounding states as a buffer.
To see how much of an effect this has, we can compare those overall calculations with ones made
just within the focal region (here, the full set of states versus just Virginia).

```{r}
buffer <- data.frame(
  # including the full set of states (VA and surrounding)
  full = baseline[population$state == "VA"],
  # including just Virginia
  none = catchment_ratio(
    population[population$state == "VA", ], hospitals[hospitals$STATE == "VA", ], traveltimes,
    weight = "gaussian", scale = 18, normalize_weight = TRUE, return_type = 1e5,
    consumers_value = "population", providers_id = "ID", providers_value = "BEDS"
  )
)
```

Overall, calculations made with routed and straight costs have a Person's _r_ of `r round(cor(buffer$full, buffer$none), 3)`,
though this is fairly sensitive to the weight function.

Looking specifically at our comparison counties...

```{r, result='asis'}
buffer_split <- get_focal_subset(buffer, block_groups$geoid)
summary_contrast(buffer_split)
```

```{r}
map_block_groups |>
  addControl("Difference in Hospital Beds Per 100K People (Full Buffer - No Buffer)", "topright") |>
  addLegend("bottomright", buffer_split$pal, buffer_split$data$difference, opacity = 1) |>
  addPolygons(
    fillColor = buffer_split$pal(buffer_split$data$difference), fillOpacity = 1, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Beds",
    label = paste0("Full - None: ", round(buffer_split$data$difference, 3))
  )
```

Higher values (blue) mean ratios are higher when the full buffer is used.
As is to be expected, we see more differences along the state border.

```{r}
splot(
  full ~ none * County, buffer_split$data,
  colors = plot_colors,
  title = FALSE, laby = "Including Surrounding States", labx = "Only Virginia"
)
```

Ratios are very similar in Loudoun, but many block groups in Fairfax have higher ratios
due to their proximity to D.C. and Maryland hospitals.

# Weight

## Norm Adjustment

One aspect of weight is the step added to make the 3-step floating catchment area: Normalization
adjustment. In this step, weights are multiplied by weights that have been normalized.
Normalized weights sum to 1 across a consumer location's connections, so multiplying by these
dampens multiple connections, which works against the priority given to highly interconnected
consumer locations.

```{r}
steps <- data.frame(
  # with adjusted weights
  three = baseline,
  # without adjusted weights
  two = catchment_ratio(
    population, hospitals, traveltimes,
    weight = "gaussian", scale = 18, return_type = 1e5,
    consumers_value = "population", providers_id = "ID", providers_value = "BEDS"
  )
)
```

Overall, calculations made with and without norm adjustments have a Person's _r_ of `r round(cor(steps$three, steps$two), 3)`.

Looking specifically at our comparison counties...

```{r, result='asis'}
step_split <- get_focal_subset(steps, block_groups$geoid)
summary_contrast(step_split)
```

```{r}
map_block_groups |>
  addControl("Difference in Hospital Beds Per 100K People (3-Step - 2-Step)", "topright") |>
  addLegend("bottomright", step_split$pal, step_split$data$difference, opacity = 1) |>
  addPolygons(
    fillColor = step_split$pal(step_split$data$difference), fillOpacity = 1, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Beds",
    label = paste0("3-Step - 2-Step: ", round(step_split$data$difference, 3))
  )
```

Higher values (blue) mean ratios are higher when weights have been adjusted.
As expected, adjusting weights with their normalized versions shifts beds toward less
interconnected regions, which particularly affects the more remote parts of Loudoun county.

```{r}
splot(
  three ~ two * County, step_split$data,
  colors = plot_colors,
  title = FALSE, laby = "With Norm-Adjusted Weights", labx = "With Original Weights"
)
```

## Scale

The weight's scale defines the size and vertical shape of the catchment areas.
For example, we can compare the baseline scale (18) with a slightly bigger scale (25) across
values representing travel times in minutes:

```{r}
# just to illustrate
Travel_Time <- 1:70
Weight <- cbind(
  catchment_weight(Travel_Time, "gaussian", scale = 18),
  catchment_weight(Travel_Time, "gaussian", scale = 25)
)
colnames(Weight) <- c(18, 25)
splot(
  Weight ~ Travel_Time,
  points = FALSE, line = "con", colors = plot_colors,
  note = FALSE, myl = c(-.1, 1), title = FALSE, leg.title = "Scale", add = {
    steps <- c(10, 20, 30, 60)
    adj <- c(-.07, .07)
    for (i in 1:2) {
      values <- cdat$`.^^.`[[i]]$y[steps]
      points(values ~ steps, col = plot_colors[i])
      text(steps, values + adj[i], prettyNum(values, digits = 2))
    }
  }
)
```

```{r}
scale <- data.frame(
  # with a weight scale of 18
  smaller = baseline,
  # with a weight scale of 25
  larger = catchment_ratio(
    population, hospitals, traveltimes,
    weight = "gaussian", scale = 25, normalize_weight = TRUE, return_type = 1e5,
    consumers_value = "population", providers_id = "ID", providers_value = "BEDS"
  )
)
```

Overall, calculations made with smaller and larger scales have a Person's _r_ of `r round(cor(scale$smaller, scale$larger), 3)`.

Looking specifically at our comparison counties...

```{r, result='asis'}
scale_split <- get_focal_subset(scale, block_groups$geoid)
summary_contrast(scale_split)
```

```{r}
map_block_groups |>
  addControl("Difference in Hospital Beds Per 100K People (Smaller - Larger)", "topright") |>
  addLegend("bottomright", scale_split$pal, scale_split$data$difference, opacity = 1) |>
  addPolygons(
    fillColor = scale_split$pal(scale_split$data$difference), fillOpacity = 1, weight = 1, color = "#000",
    highlightOptions = highlightOptions(color = "#fff"), group = "Beds",
    label = paste0("Smaller - Larger: ", round(scale_split$data$difference, 3))
  )
```

Higher values (blue) mean ratios are higher with the smaller scale.
A larger scale works to more evenly distribute beds, generally shifting them from nearer to farther
consumer locations.

```{r}
splot(
  smaller ~ larger * County, scale_split$data,
  colors = plot_colors,
  title = FALSE, laby = "With A Smaller Scale (18)", labx = "With A Larger Scale (25)"
)
```

This is incidentally very similar to the effect of norm adjusting weights because consumer
locations with more provider connections also tend to be closer to those providers.
